---
title: "covid 19"
author: "MarenaO"
date: "Spring 2022 Semester"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>% mutate(across(where(is.character), as.factor))
```


### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

According to the barplot,I can see that there are not many people that are dying. Although, one barplot shows that of the people that are dying, it looks like more males are dying than females. This amount looks to be about twice as many males dying than females.
  





  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

According to the numerical results, 1.4% of females have died and 3% of males have died. This amount of males that died is more than twice the amount of females that died.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

According to the inferential results, it is evident that 25 females have died and that 44 males have died. The p-value was 0.0012, which is a low p-value. Therefore, I can conclude that there was a significant difference between the males dying and the females dying. The Fisher test shows that the odds of dying of male vs. female is 2.3 to 1, which is a very significant difference. 

## Conclusions

